SUPPORTING INFORMATION: Towards atomistic resolution structure of phosphatidylcholine headgroup and glycerol backbone at different ambient conditions\footnote{Publication about results presented in the NMRlipids project}

### 2015-08-22 HGmodelMANUSCRIPT/HGmodel_ACSsupplementary.tex
Added description of FF ranking. Added S in front of figure numbers. Fixed some typos.

@@

@@

\section{Author Contributions}
{\it Alexandru Botan} provided simulation results for charmm36-UA\\
{\it Fernando Favela} prepared, performed and analyzed simulations
with popc and cholesterol for the CHARMM36 All-Atom FF. \\
{\it Patrick F. J. Fuchs} Ran and analyzed the Poger simulations. Provided scientific information which significantly advanced the project (signs and forking of order parameters). \\
{\it Matti Javanainen} prepared and performed simulations with multiple lipid models and analyzed the results. Supervised the work of JT.\\
{\it Matej Kanduc} provided simulation results for the Berger DLPC model \\
{\it Waldemar Kulig} prepared the MD simulations with cholesterol for the MacRog FF \\
{\it Antti Lamberg} Prepared and performed simulations with Berger and its variants to show the importance of the signs and stereospecific labeling of the order parameters. \\
{\it Claire Loison} provided simulation results for charmm36-UA. \\
{\it Alexander Lyubartsev} provided simulation results for Högberg08 force field.\\
{\it Markus S. Miettinen} Co-designed the project with OHSO and supported in the work management. Provided trajectories for Berger DMPC; prepared parameter files for 72 lipids \\
{\it Luca Monticelli} Critical discussions in all phases of the project. Collaboration with Jukka  M{\"a}{\"a}tt{\"a}. \\
{\it Jukka M{\"a}{\"a}tt{\"a}}  prepared and performed simulations with Berger and Slipids models and analyzed the results.\\
{\it O. H. Samuli Ollila} Co-designed the project with MSM and managed the work. Ran and analyzed several simulations. Wrote the manuscript. \\
{\it Marius Retegan} Prepared and validated the GROMACS compatible parameter files for GAFFlipid and Lipid14 force fields.\\
{\it Tomasz Rog} Provided the topologies for the MacRog model and also the full hydration simulation data with this model. \\
{\it Hubert Santuz} prepared and performed the cholesterol simulations with CHARMM36 and analyzed the results. 
Made significant contribution in the data management of the project.\\
{\it Joona Tynkkynen} prepared and performed the dehydration simulations with the MacRog FF.\\

>>>>

\renewcommand{\thefigure}{S\arabic{figure}}

@@

\section{Details of the rough subjective force field ranking (Figure 4)}

The assessment was based fully on the Fig. 2.
%
First, for each carbon (the columns in Fig. 4) in each force field (the rows),
we looked separately at deviations in magnitude and forking.

{\bf Magnitude} deviations, i.e., how close to the experimentally obtained C--H order parameters (OPs)
the force-field-produced OPs were.
%
For each carbon, the following 5-step scale was used:
%
\begin{description}
\item [0 (~):] \noindent {More than half of all the calculated OPs (that is, of all different hydrogens in all different lipids) were within the {\it subjective sweet spots} (SSP, blue-shaded areas in Fig. 2).}
%
\item [1 ({\textsf{\tiny M}}):] \noindent {All the calculated OPs  were $< 0.03$ units away from the SSP.}
%
\item [2  ({\textsf{\small M}}):] \noindent {All the calculated OPs  were $< 0.05$ units away from the SSP.}
%
\item [3 ({\textsf{\large M}}):] \noindent {All the calculated OPs  were $< 0.10$ units away from the SSP.}
%
\item [4 ({\textsf{\Large M}}):] \noindent {Some of the calculated OPs  were $> 0.10$ units  away from the SSP.}
\end{description}

{\bf Forking} deviations, i.e., how well the difference in order parameters of two hydrogens attached to a given carbon matched that obtained experimentally. Note that this is not relevant for $\mathrm{g_2}$, which has only one hydrogen. For the $\beta$ and $\alpha$ carbons, for which no forking is experimentally seen, the following 5-step scale was used:
\begin{description}
\item [0 (~):] \noindent {The distance $D$ between the dots (that mark the measurement-time-weighted averages in Fig.~2) was $< 0.005$ units for all the calculated OPs (that is, for all different lipids).}
%
\item [1 ({\textsf{\tiny F}}):] \noindent {$D < 0.015$.}
%
\item [2  ({\textsf{\small F}}):] \noindent {$ 0.015 < D < 0.030$.}
%
\item [3 ({\textsf{\large F}}):] \noindent {$ 0.030 < D < 0.050$.}
%
\item [4 ({\textsf{\Large F}}):] \noindent {$ 0.050 < D $.}
\end{description}
%
For the $\mathrm{g_3}$ carbon, for which a small forking of 0.02 is indicated by experiments, the following 5-step scale was used:
%
\begin{description}
\item [0 (~):] \noindent {$0.010 < D< 0.030$.}
%
\item [1 ({\textsf{\tiny F}}):] \noindent {$(D <0.010)$  OR $(0.030 < D < 0.040)$.}
%
\item [2  ({\textsf{\small F}}):] \noindent {$0.040 < D < 0.055$.}
%
\item [3 ({\textsf{\large F}}):] \noindent {$0.055 < D < 0.070$.}
%
\item [4 ({\textsf{\Large F}}):] \noindent {$0.070 < D$.}
\end{description}
%
For the $\mathrm{g_1}$ carbon, for which a considerable forking of 0.15 is experimentally seen, the following 5-step scale was used:
%
\begin{description}
\item [0 (~):] \noindent {$0.13 < D < 0.17$.}
%
\item [1 ({\textsf{\tiny F}}):] \noindent {$(0.11 < D < 0.13)$ OR $(0.17 < D < 0.19)$.}
%
\item [2  ({\textsf{\small F}}):] \noindent {$(0.09 < D < 0.11)$ OR $(0.19 < D < 0.21)$.}
%
\item [3 ({\textsf{\large F}}):] \noindent {$(0.07 < D < 0.09)$ OR $(0.21 < D < 0.23)$.}
%
\item [4 ({\textsf{\Large F}}):] \noindent {$(D<0.07)$ OR $(0.23<D)$.}
\end{description}

Based on these assessments of magnitude and forking deviations,
each carbon was then assigned to one of the following groups:
"within experimental error"
(magnitude and forking deviations both on step 0 of the scales described above),
"almost within experimental error"
(sum of the magnitude and forking deviation steps 1 or 2),
"clear deviation from experiments"
(sum of magnitude and forking deviation steps from 3 to 5), and
"major deviation from experiments"
(sum of magnitude and forking deviation steps from 6 to 8).
These groups are indicated by colors in Fig.~2.
(Note that for $\mathrm{g_2}$, for which there can be no forking,
the corresponding group assigment limits were: 0, 1, 2, and 3.)

Finally, the total ability of the force field to describe the headgroup and
glycerol structure was estimated.
To this end, the groups were given the following weights:
0 (within experimental error),
1 (almost within experimental error),
2 (clear deviation from experiments),
4 (major deviation from experiments),
and the weights of the five carbons were summed up.
The sum, given in the $\Sigma$-column of Fig.~2,
was then used to (roughly and subjectively, as should be clear from the
above description) rank the force fields.

@@

\section{Author Contributions}
{\it Alexandru Botan} provided simulation results for charmm36-UA\\
{\it Fernando Favela} prepared, performed and analyzed simulations
with popc and cholesterol for the CHARMM36 All-Atom FF. \\
{\it Patrick F. J. Fuchs} ran and analyzed the Poger simulations. Provided scientific information which significantly advanced the project (signs and forking of order parameters). \\
{\it Matti Javanainen} prepared and performed simulations with multiple lipid models and analyzed the results. Supervised the work of JT.\\
{\it Matej Kanduc} provided simulation results for the Berger DLPC model \\
{\it Waldemar Kulig} prepared the MD simulations with cholesterol for the MacRog FF \\
{\it Antti Lamberg} prepared and performed simulations with Berger and its variants to show the importance of the signs and stereospecific labeling of the order parameters. \\
{\it Claire Loison} provided simulation results for charmm36-UA. \\
{\it Alexander Lyubartsev} provided simulation results for Högberg08 force field.\\
{\it Markus S. Miettinen} co-designed the project with OHSO and supported in the work management. Provided Berger DMPC trajectory; prepared parameter files for 72 lipids %POPC \& DPPC
CHARMM36. Made Figures 2--4 and 7--9.\\
{\it Luca Monticelli}: Critical discussions in all phases of the project. Collaboration with Jukka  M{\"a}{\"a}tt{\"a}. \\
{\it Jukka M{\"a}{\"a}tt{\"a}}  prepared and performed simulations with Berger and Slipids models and analyzed the results.\\
{\it O. H. Samuli Ollila} co-designed the project with MSM and managed the work. Ran and analyzed several simulations. Wrote the manuscript. \\
{\it Marius Retegan} prepared and validated the GROMACS compatible parameter files for GAFFlipid and Lipid14 force fields.\\
{\it Tomasz Rog} provided the topologies for the MacRog model and also the full hydration simulation data with this model. \\
{\it Hubert Santuz} prepared and performed the cholesterol simulations with CHARMM36 and analyzed the results. 
Made significant contribution in the data management of the project.\\
{\it Joona Tynkkynen} prepared and performed the dehydration simulations with the MacRog FF.\\
